Data profiling project - Saheli Kar
========================================================

## Part 1: 
I am choosing a data set of Real Estate. The csv file is shared in the GitHub.

```{r}
RealEstateDf<-data1<- read.csv(file.choose(), header=TRUE)
head(RealEstateDf)
```

## Part 2:

### The variables in the Sacramento Real Estate data sets are: 
<b>street - </b>  The street address of the property. <br/>
<b>city - </b> The city where the property is located. <br/>
<b>zip - </b> The zip code of the property. <br/>
<b>state - </b>The state where the property is located. <br/>
<b>beds - </b> The number of bedrooms in the property. <br/>
<b>baths - </b> The number of bathrooms in the property. <br/>
<b>sq__ft - </b> The area of the property in square feet. <br/>
<b>type - </b> The type of the property whether it is Residential/Condominium/ Multifamily <br/>
<b>sale_date - </b> The date when the property is sold. <br/>
<b>price - </b> The price of the propert sold. <br/>
<b>latitude - </b> The latitude loaction of the property.<br/>
<b>longitude - </b> The longitude location of the property. <br/>

### The correaltion of the variables are provided with ggpairs()

```{r fig.width=20, fig.height=20}

#install.package(psych)
#require(psych)
#pairs(~zip+beds+baths+sq__ft+type+sale_date+price, data=RealEstateDf)

install.packages("GGally")
require(GGally)

ggpairs(data=RealEstateDf, columns=c("zip", "beds", "baths", "sq__ft", "type", "sale_date", "price"))

```

Above we can see the correlation between the variables:<br/>
zip, beds, baths, sq__ft, type, sale_date, price<br/><br.>

Interesting correlation to be noted:
* Most of the 2 and 4 bedrooms property are sold in the areas around the zipcode 95800. Most of the 3 bedrooms are sold in the area with zip code 95600. Only one area had 8 bedrooms on sale.
* Very intersting positive correlation between beds and baths found.
* Also very positive correlation found between sqft and baths; and sqft and beds
* Most of the inexpensive property sold in the area around 95800. While Other areas sold realtively more exepnsive properties than the area under 95800.
* Most of the properties sold are Residential and most properties are inexpensive.
* Price and sq__ft doesnt hold a spreaded correlation.

### Recoding and binning:
Recoding the price and the sq__ft variable.

```{r}
RealEstateDf$priceCategory[RealEstateDf$price<=100000]<-"Inexpensive"
RealEstateDf$priceCategory[RealEstateDf$price>100000 & RealEstateDf$price<=200000]<-"AveragePriced"
RealEstateDf$priceCategory[RealEstateDf$price>200000]<-"Expensive"

RealEstateDf$AreaCategory[RealEstateDf$sq__ft<1000]<-"Smaller"
RealEstateDf$AreaCategory[RealEstateDf$sq__ft>1000 & RealEstateDf$sq__ft<2000] <-"Bigger"
RealEstateDf$AreaCategory[RealEstateDf$sq__ft>2000]<-"Very Big"

head(RealEstateDf, n=50)

```

### Frequency of the Price Category with respect to the cities
```{r}
require(dplyr)

priceCategoryFreq<-select(RealEstateDf, city, priceCategory) %>%
group_by(city, priceCategory) %>%
summarise(count=n()) 

show(priceCategoryFreq)

```

### Frequency of the Area Category with respect to the cities
```{r}
require(dplyr)

areaCategoryFreq<-select(RealEstateDf, city, AreaCategory) %>%
group_by(city,AreaCategory) %>%
summarise(count=n())

show(areaCategoryFreq)

```

### Distribution of property prices with respect to longitude and latitude
This will help in locating area which has less priced propert and highly priced property
```{r}
require(ggplot2)
require(reshape)
require(gridExtra)

longitudeMap<-ggplot(data = RealEstateDf) + geom_point(aes(x=longitude, y=priceCategory ))+ ggtitle("Distribution of property prices with respect to longitude")

latitudeMap<-ggplot(data = RealEstateDf) + geom_point(aes(x=latitude, y=priceCategory ))+ ggtitle("Distribution of property prices with respect to latitude")

grid.arrange(longitudeMap,latitudeMap, ncol=1 )
```