# Week 5 Assignment by Saheli Kar
========================================================
## Answer 1
1. How many voters are over 25?
2. How many voters voted prefferred Partan bree?
3. What are the average number of voters from both the towns who preffered Cullen skink?


## Answer 2
```{r}
require(dplyr)
preference<-c("Cullen skink", "Partan bree")

Edinburg<-data.frame(c(80100,35900 ), c(143000, 214800))

Glasgow<-data.frame(c(99400, 43000), c(150400, 207000))

messyData<-data.frame(preference, Edinburg, Glasgow)

names(messyData)<-c("preference", "Edinburg.16to24", "Edinburg.25plus", "Glasgow.16to24", "Glasgow.25plus")

show(messyData)

```

## Answer 3

```{r}
require(dplyr)
require(tidyr)

tidyData<-messyData %>%
gather(key, value, Edinburg.16to24, Edinburg.25plus, Glasgow.16to24, Glasgow.25plus) %>%
separate(key, c("City", "Age"))

show(tidyData)
```

## Answer 4
```{r}
AllVotersOver25<-aggregate(value~Age, tidyData, sum) %>%
subset(Age=="25plus", select=value)
show(AllVotersOver25)

prefferedParatnBree<-aggregate(value~preference, tidyData, sum) %>%
subset(preference=="Partan bree", select=value)
show(prefferedParatnBree)

avgVoterCullenSkink<-aggregate(value~preference, tidyData, sum) %>%
subset(preference=="Cullen skink", select=value)/length(unique(tidyData$City)) 
show(avgVoterCullenSkink)

```

## Answer 5
Yes I can think of more questions from the datasets like:

1. Show the list of count all voters from both the towns who are within the age 16-24
2. Select the count of all the voters from Edinburg who voted for Partan bree and age is 25+
3. We can build many more questions like this.